---
title: "Support Vector Machine"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: espresso
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(rio)
library(colorblindr)
library(gghighlight)
library(forcats)
library(ggrepel)
library(knitr)
library(kableExtra)
library(reactable)
library(plotly)
library(glue)
library(fs)
library(rstatix)
library(ggpubr)
library(writexl) 
library(remotes)
library(profvis) 
library(tidymodels)
library(tune)
library(glmnet)
library(baguette)
library(parsnip)
library(doParallel)
library(vip)
library(pdp)
library(patchwork)
library(ranger)
library(future)
library(rio)
library(bit64)
library(rpart.plot)
library("e1071")
library(GGally)


theme_set(theme_minimal(15) +
            theme(legend.position = "bottom",
                  panel.grid.major.x = element_line(color = "gray60"),
                  panel.grid.minor.x = element_blank(),
                  panel.grid.major.y = element_blank())
          )
```

```{r global, include=FALSE}
#all clean sims data
sims_concussion_data <- read_csv(here("data", "sims_concussion_data.csv"))

sims_concussion_data <- sims_concussion_data %>% 
  mutate(age = as.factor(age))

simsimp <- read_csv(here("data", "clean_impact_sims_data.csv"))

mod_data <- read_csv(here("data", "clean_impact_sims_data.csv"))

mod_datav2 <- read_csv(here("data", "clean_impact_sims_data.csv"))

str(simsimp)

simsimp <- simsimp %>% 
  mutate(dataset = as.factor(dataset),
         school_year = as.factor(school_year),
         school = as.factor(school),
         league = as.factor(league),
         gender = as.factor(gender),
         age = as.factor(age),
         sport = as.factor(sport),
         injury = as.factor(injury)) %>% 
  mutate_if(is.numeric, round, digits = 3)

mod_data <- mod_data %>% 
  mutate(dataset = as.factor(dataset),
         school_year = as.factor(school_year),
         school = as.factor(school),
         league = as.factor(league),
         gender = as.factor(gender),
         sport = as.factor(sport),
         injury = as.factor(injury)) %>% 
  mutate_if(is.numeric, round, digits = 3)

str(mod_data)

mod_datav2 <- mod_datav2 %>% 
  mutate(dataset = as.factor(dataset),
         school_year = as.factor(school_year),
         school = as.factor(school),
         league = as.factor(league),
         gender = as.factor(gender),
         sport = as.factor(sport),
         injury = as.factor(injury)) %>% 
  mutate_if(is.numeric, round, digits = 3)

str(mod_datav2)

```

```{r, include=FALSE}
# creating new data set for machine learning models

names(simsimp)
str(simsimp)
names(mod_data)

mod_data[mod_data$age <= 16, "age_group"] <- "13-16"
mod_data[mod_data$age >=17, "age_group"] <- "17-18"

names(mod_data)

mod_data %>% 
  count(age_group)

str(mod_data)

mod_data <- mod_data %>% 
  mutate(age_group = as.factor(age_group))

str(mod_data)

mod_data[mod_data$dys_btwn_onset_rtp_3 <= 6, "rtl_group"] <- "0-6"
mod_data[mod_data$dys_btwn_onset_rtp_3 >= 7 & mod_data$dys_btwn_onset_rtp_3 <= 12,
         "rtl_group"] <- "7-12"
mod_data[mod_data$dys_btwn_onset_rtp_3 >= 13 & mod_data$dys_btwn_onset_rtp_3 <= 18,
         "rtl_group"] <- "13-18"
mod_data[mod_data$dys_btwn_onset_rtp_3 >= 19 & mod_data$dys_btwn_onset_rtp_3 <= 24,
         "rtl_group"] <- "19-24"
mod_data[mod_data$dys_btwn_onset_rtp_3 >= 25 & mod_data$dys_btwn_onset_rtp_3 <= 30,
         "rtl_group"] <- "25-30"

str(mod_data)

mod_data <- mod_data %>% 
  mutate(rtl_group = as.factor(rtl_group))

mod_data %>% 
  count(rtl_group)

mod_data_2 <- mod_data

mod_data_2[mod_data_2$total_symptom_score_post_injury_1 <= 16,
           "test_1_pcss_group"] <- "0-16"
mod_data_2[mod_data_2$total_symptom_score_post_injury_1 >= 17 &
           mod_data_2$total_symptom_score_post_injury_1 <= 32,
           "test_1_pcss_group"] <- "17-32"
mod_data_2[mod_data_2$total_symptom_score_post_injury_1 >= 33 &
           mod_data_2$total_symptom_score_post_injury_1 <= 48,
           "test_1_pcss_group"] <- "33-48"
mod_data_2[mod_data_2$total_symptom_score_post_injury_1 >= 49 &
           mod_data_2$total_symptom_score_post_injury_1 <= 64,
           "test_1_pcss_group"] <- "49-64"
mod_data_2[mod_data_2$total_symptom_score_post_injury_1 >= 65 &
           mod_data_2$total_symptom_score_post_injury_1 <= 80,
           "test_1_pcss_group"] <- "65-80"
mod_data_2[mod_data_2$total_symptom_score_post_injury_1 >= 81,
           "test_1_pcss_group"] <- "81 or higher"

str(mod_data_2)

mod_data_2 <- mod_data_2 %>% 
  mutate(test_1_pcss_group = as.factor(test_1_pcss_group))

str(mod_data_2)

mod_data_2 %>% 
  count(test_1_pcss_group)


# per Tom's suggestion, divding RTL group into two levels - splitting around 12, which is the mean


mod_datav2[mod_datav2$age <= 16, "age_group"] <- "13-16"
mod_datav2[mod_datav2$age >=17, "age_group"] <- "17-18"

mod_datav2[mod_datav2$dys_btwn_onset_rtp_3 <= 12, "rtl_group"] <- "0-12"
mod_datav2[mod_datav2$dys_btwn_onset_rtp_3 >= 13, "rtl_group"] <- "13-30"

mod_datav2[mod_datav2$total_symptom_score_post_injury_1 <= 16,
           "test_1_pcss_group"] <- "0-16"
mod_datav2[mod_datav2$total_symptom_score_post_injury_1 >= 17 &
           mod_datav2$total_symptom_score_post_injury_1 <= 32,
           "test_1_pcss_group"] <- "17-32"
mod_datav2[mod_datav2$total_symptom_score_post_injury_1 >= 33 &
           mod_datav2$total_symptom_score_post_injury_1 <= 48,
           "test_1_pcss_group"] <- "33-48"
mod_datav2[mod_datav2$total_symptom_score_post_injury_1 >= 49 &
           mod_datav2$total_symptom_score_post_injury_1 <= 64,
           "test_1_pcss_group"] <- "49-64"
mod_datav2[mod_datav2$total_symptom_score_post_injury_1 >= 65 &
           mod_datav2$total_symptom_score_post_injury_1 <= 80,
           "test_1_pcss_group"] <- "65-80"
mod_datav2[mod_datav2$total_symptom_score_post_injury_1 >= 81,
           "test_1_pcss_group"] <- "81 or higher"


str(mod_datav2)

mod_datav2 <- mod_datav2 %>% 
  mutate(rtl_group = as.factor(rtl_group),
         test_1_pcss_group = as.factor(test_1_pcss_group),
         age_group = as.factor(age_group))

str(mod_datav2)

mod_datav2 %>% 
  count(rtl_group)
```


```{r split, include=FALSE}

svm_data <- mod_data_2 %>% 
  select(gender, age_group, test_1_pcss_group, rtl_group)

str(svm_data)

# data where rtl_group is split into two levels only 

svm_data2 <- mod_datav2 %>% 
  select(gender, age_group, test_1_pcss_group, rtl_group)

str(svm_data2)

```


## Model with Five RTL Group Levels {.tabset .tabset-fade .tabset-pills}


```{r, include=FALSE}
# other SVM example code 

set.seed(3)
svm_split <- initial_split(svm_data, strata = rtl_group)


set.seed(3)
svm_train <- training(svm_split)
svm_test <- testing(svm_split)

svm_model <- svm(rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_train,
              kernel = "radial")

svm_model

```

### Initial Model Results

```{r, include=TRUE}
table(Prediction = predict(svm_model, svm_train), Actual = svm_train$rtl_group)
```

### Tuned Model Accuracy 

Tuning the model generates an accuracy of 0.606

```{r, include=FALSE}
tuned.svm <- tune.svm(rtl_group ~ gender + age_group + test_1_pcss_group,
                      data = svm_train, kernel = "radial",
        gamma = seq(1/2^nrow(svm_data),1, .01), cost = 2^seq(-6, 4, 2))
```

```{r, include=TRUE}
tuned.svm
```

### Best Model Parameters 

```{r, include=FALSE}
best.svm <- best.svm(rtl_group ~ gender + age_group + test_1_pcss_group,
                     data = svm_train, kernel = "radial")

```

```{r, include=TRUE}
best.svm
```

### Best Model Predictions

```{r, include=TRUE}
best.svm.pred <- predict(best.svm, svm_test)
table(Prediction = best.svm.pred, Actual = svm_test$rtl_group)
```

### Best Model Accuracy 

The generated accuracy does not seem realistic or accurate. 

```{r, include=TRUE}
sum(svm_test$rtl_group == best.svm.pred)/50
```

## Model with Five RTL Group Levels (Alternative Code) {.tabset .tabset-fade .tabset-pills}

```{r, include=FALSE}
svm_modelb <- svm(rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_data,
              kernel = "radial")
```


```{r, include=FALSE}
ggpairs(svm_data, ggplot2::aes(colour = rtl_group, alpha = 0.4))
```


```{r, include=FALSE}
pred = predict(svm_modelb, svm_data)

tab = table(Predicted=pred, Actual = svm_data$rtl_group)


1-sum(diag(tab)/sum(tab))
```

### Model Predictions

```{r, include=TRUE}
tab
```

### Model Accuracy

This code generates an accuracy of 0.624

```{r, include=TRUE}
1-sum(diag(tab)/sum(tab))
```

### Parameter Tuning Plot

```{r, include=FALSE}
set.seed(123)
tmodel = tune(svm, rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_data,
     ranges=list(epsilon= seq(0,1,0.1), cost = 2^(2:7)))

```

```{r, include=TRUE}
plot(tmodel)
```

### Parameter Tuning Summary 

```{r, include=TRUE}
summary(tmodel)
```

### Summary of Best Model

```{r, include=TRUE}
mymodel=tmodel$best.model
summary(mymodel)
```

### Final Prediction Table

```{r, include=FALSE}
pred1 = predict(mymodel, svm_data)
tab1 = table(Predicted=pred1, Actual = svm_data$rtl_group)
```

```{r, include=TRUE}
tab1
```

### Final Prediction Accuracy

The model generates an accuracy of 0.587

```{r, include=TRUE}
1-sum(diag(tab1)/sum(tab1))
```








## Model with Two RTL Group Levels {.tabset .tabset-fade .tabset-pills}


```{r, include=FALSE}
# other SVM example code 

set.seed(3)
svm_split2 <- initial_split(svm_data2, strata = rtl_group)


set.seed(3)
svm_train2 <- training(svm_split2)
svm_test2 <- testing(svm_split2)

svm_model2 <- svm(rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_train2,
              kernel = "radial")

svm_model2

```

### Initial Model Results

```{r, include=TRUE}
table(Prediction = predict(svm_model2, svm_train2), Actual = svm_train2$rtl_group)
```

### Tuned Model Accuracy 

Tuning the model generates an accuracy of 0.398

```{r, include=FALSE}
tuned.svm2 <- tune.svm(rtl_group ~ gender + age_group + test_1_pcss_group,
                      data = svm_train2, kernel = "radial",
        gamma = seq(1/2^nrow(svm_data2),1, .01), cost = 2^seq(-6, 4, 2))
```

```{r, include=TRUE}
tuned.svm2
```

### Best Model Parameters 

```{r, include=FALSE}
best.svm2 <- best.svm(rtl_group ~ gender + age_group + test_1_pcss_group,
                     data = svm_train2, kernel = "radial")

```

```{r, include=TRUE}
best.svm2
```

### Best Model Predictions

```{r, include=TRUE}
best.svm.pred2 <- predict(best.svm2, svm_test2)
table(Prediction = best.svm.pred2, Actual = svm_test2$rtl_group)
```

### Best Model Accuracy 

The generated accuracy does not seem realistic or accurate. 

```{r, include=TRUE}
sum(svm_test2$rtl_group == best.svm.pred2)/50
```

## Model with Two RTL Group Levels (Alternative Code) {.tabset .tabset-fade .tabset-pills}

```{r, include=FALSE}
svm_model2b <- svm(rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_data2,
              kernel = "radial")
```


```{r, include=FALSE}
ggplotly(ggpairs(svm_data2, ggplot2::aes(colour = rtl_group, alpha = 0.4)))
```


```{r, include=FALSE}
pred2 = predict(svm_model2b, svm_data2)

tab2 = table(Predicted=pred2, Actual = svm_data2$rtl_group)


1-sum(diag(tab2)/sum(tab2))
```

### Model Predictions

```{r, include=TRUE}
tab2
```

### Model Accuracy

This code generates an accuracy of 0.405

```{r, include=TRUE}
1-sum(diag(tab2)/sum(tab2))
```

### Parameter Tuning Plot

```{r, include=FALSE}
set.seed(123)
tmodel2 = tune(svm, rtl_group ~ gender + age_group + test_1_pcss_group, 
              data = svm_data2,
     ranges=list(epsilon= seq(0,1,0.1), cost = 2^(2:7)))

```

```{r, include=TRUE}
plot(tmodel2)
```

### Parameter Tuning Summary 

```{r, include=TRUE}
summary(tmodel2)
```

### Summary of Best Model

```{r, include=TRUE}
mymodel2=tmodel2$best.model
summary(mymodel2)
```

### Final Prediction Table

```{r, include=FALSE}
pred1b = predict(mymodel2, svm_data2)
tab1b = table(Predicted=pred1b, Actual = svm_data2$rtl_group)
```

```{r, include=TRUE}
tab1b
```

### Final Prediction Accuracy

The model generates an accuracy of 0.381

```{r, include=TRUE}
1-sum(diag(tab1b)/sum(tab1b))
```